\singlespacing{\textbf{Efficient Pollution Abatement in Electricity Markets with Intermittent Renewable Energy}}\thanks{We would like to thank Prof. Roger Klein for helpful comments and the Sun Grant (award number 122682) for funding this project.}

### 2019-06-21 documents/draft.tex
Update draft
- add contact info
- fix abstract key words and jel
- move references
- move figures and tables
- minor fixes

\documentclass[11pt,a4paper]{extarticle}
\usepackage[margin=1in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{booktabs} % for toprule, midrule and bottomrule
\usepackage{adjustbox}
\usepackage{amsmath}
\usepackage{bbold}
\usepackage{etoolbox}
\usepackage{setspace} % for \onehalfspacing and \singlespacing macros
\usepackage[hidelinks]{hyperref}
\usepackage{array}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{caption}
\usepackage{pdflscape}
\usepackage{caption}
\usepackage{tabularx}
\usepackage{authblk}
\usepackage{float}
\usepackage{siunitx}
\usepackage{titlesec}
\usepackage{pgfplots}
\usepackage[authoryear]{natbib}
\usepackage{scrextend}
\usepackage{nicefrac}
\usepackage{enumitem}
\usepackage{multirow}
\usepackage{xcolor}
\usepackage{cleveref}
\usepackage{varwidth}
\usepackage{gensymb}
%\usepackage{showframe}
%\usepackage{lipsum}

% set space
%\doublespacing

@@

\titleformat{\section}
{\bf\centering\large}{\thesection}{1em}{}

\titleformat{\subsection}
{\itshape\centering}{\thesubsection}{1em}{}

\titleformat{\subsubsection}
{\bf}{\thesubsubsection}{1em}{}

% unicode chars for plots
\DeclareUnicodeCharacter{2212}{$-$}

% booktabs
\setlength\heavyrulewidth{0.06em} % 0.01em> midrule

% images
%\graphicspath{ {D:/Users/saketh/Documents/GitHub/Energy-Intermittency-Paper/figures/} }

@@

% array
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

@@

% sig stars
\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}

@@

	\title{\singlespacing{\textbf{Efficient Pollution Abatement in Electricity Markets with Intermittent Renewable Energy}}\thanks{We would like to thank Prof. Roger Klein for helpful comments and the Sun Grant (award number 122682) for funding this project.}}

	\author[]{Saketh Aleti \, and Gal Hochman}

	\affil[]{Rutgers University}

@@

	% Syntax: \begin{addmargin}[<left indentation>]{<indentation>}
	\begin{addmargin}[0.5in]{0.5in}
		\textit{In this paper, we present a model of the electricity sector where generation technologies are intermittent. The economic value of a electricity generation technology is given by integrating its production profile with the market price of electricity. We parametrize this model empirically and generate numerical results. The introduction of intermittency results in a non-constant elasticity of substitution between renewable and fossil energy. Consequently, this suggests that the efficacy and welfare effects of carbon taxes and renewable subsidies vary geographically. Subsidizing research into battery technology can mitigate this distributional side effect while complementing traditional policies used to promote renewable energy. (JEL Q28, Q41, Q48, Q52, Q55, Q58)
		}
	\end{addmargin}

@@

	Consumers purchase a quantity of electricity $Z_t$ in each period $t$. Furthermore, they demand a greater quantity of electricity in certain periods; for instance, consumers need more electricity during the middle of the day more than at night. At the same time, consumers are willing to shift their consumption from one period to another in response to a shift in prices. Overall, these characteristics can be captured using a standard CES utility function.  Furthermore, since CES preferences are homothetic, we may aggregate the consumers into a single representative consumer. This consumer has the utility function
	\begin{equation}
	U = \left( \sum_t \alpha_t Z_t^\phi  \right)^{1/\phi}
	\end{equation}
	where $\sigma = 1/(1-\phi)$ is the intertemporal elasticity of substitution for electricity consumption. To ensure that indifferences surfaces are convex, we have $\phi \in (0,1)$ or, equivalently,  $\sigma > 0$; additionally, since electricity consumption increases utility, we must have $\alpha_t > 0$ for all $t$.  For simplicity, we also define $\sum_t \alpha_t = 1$, so that a 1\% increase in electricity consumption in all periods increases utility by 1\%; consequently, we then have $\alpha_t \in (0,1)$. The budget constraint is given by
	\begin{equation}
	I = \sum_t p_t Z_t
	\end{equation}
	where $p_t$ is the price of electricity in period $t$ and $I$ is the income. Our representative consumer maximizes utility against this budget constraint; the first order conditions of this problem imply:
	\begin{align}
	Z_t &= \left(\frac{\alpha_t}{p_t} \right)^\sigma \frac{I}{P} \\
	P &= \sum_t \alpha_t^\sigma p_t^{1-\sigma}
	\end{align}
	where $P$ is the price index. Note that this model naturally does not allow for blackouts in equilibrium, since the price of electricity in any period gets arbitrarily large as the quantity of energy consumed in that period approaches 0. Furthermore, note that prices must be positive; although this is sometimes violated in reality, we do not believe that this assumption significantly affects our analysis.

@@

	Next, our representative firm sets $X$ once to maximize total profit while facing the cost function $C(X_1, \dots, X_n)$.  \citet{HH} note that past literature has argued for concave cost functions in the energy market due to effects such as economies of scale and learning-by-doing; on the other hand, standard cost functions are generally convex. So, like Helm and Meir, we take an intermediate approach by using a linear cost function. Specifically, we have $C(X) \equiv \sum_i c_i X_i \equiv c^T X$ where $c_i$ is the cost per unit of $X_i$. Total profit is given by 
	\begin{equation}
	\Pi = p^T Z - c^T X
	\end{equation}
	To simplify the algebra, we set the number of technologies equal to the number of periods $(n=m)$. Additionally, we further require that the output per unit of each technology is unique and non-negative in each period; in other words, the output per unit of one technology is not a linear combination of those of the other technologies in our set. This then implies that $\xi$ is of full rank and therefore invertible.  Now, maximizing profit, we find the first order condition:
	\begin{equation}
	\frac{\partial \Pi}{\partial X} = 0 \implies p = \xi^{-1} c
	\end{equation}
	Combining this FOC with the demand equation (Equation (3)) allows us to find the equilibrium. Generally, the equilibrium results for any number of technologies ($n$) are analytic, but they are difficult to interpret due to the number of parameters involved.

@@

	Now, recall the demand equation from earlier (Equation 3)
	$$Z_t = \left(\frac{\alpha_t}{p_t} \right)^\sigma \frac{I}{P}$$
	where $P$ is the price index and $I$ is income. Retail customers pay fixed rates each month for electricity, hence $p_t$ is constant within each month; and, we expect that income $I$ does not vary significantly on a daily basis. Consequently, all variation in intramonthly demand is due to the share parameter $\alpha$ and the elasticity $\sigma$. But, this creates a problem; since retail consumers do not face prices that vary each hour, we cannot estimate  $\sigma$ on an hourly basis.

@@

	Firstly, recall the demand equation (Equation 3) from our general model:
	\begin{align*}
	Z_t &= \left(\frac{\alpha_t}{p_t} \right)^\sigma \frac{I}{P} \\
	P &= \sum_t \alpha_t^\sigma p_t^{1-\sigma}
	\end{align*}
	For any pair of electricity outputs $Z_t$ and $Z_s$, we have:
	$$\frac{Z_t}{Z_s} = \left(\frac{\alpha_t \, p_s}{\alpha_s \, p_t} \right)^\sigma \\$$
	Taking logs on both sides and letting $i$ represent different observations, we may rewrite this in a form more suitable for estimation. 
	\begin{align*}
	\ln (Z_{t, i} / Z_{ s, i}) &= -\sigma \ln (P_{t,i} / P_{s,i}) + \sigma \ln (\alpha_{t,i} / \alpha_{s,i}) 
	\end{align*}
	Our data differentiates consumption for each state in the US, so we let $i$ refer to a particular state. Additionally, most consumers pay monthly fixed rates for electricity, so we can, at most, estimate this equation on a monthly basis; hence, $t$ and $s$ refer to different months. Also, note that each observation only corresponds to a single state $i$; this is because consumers within each state can substitute consumption across time, but consumers in different states do not substitute consumption with one another.

@@

	We collect monthly data from the \cite{EIANetgen} on retail electricity prices and consumption for each state in the US from 2011 to 2018. Additionally, also from the EIA, we obtain data on the average cost of coal for electricity generation for each state and month.\footnote{ The coal price data set contains a large number of missing values due to privacy concerns; however, we do not expect that these missing values are correlated with the data itself.} We deflated both electricity and coal prices over time using the PCEP Index provided by the  \cite{USBEA}. 
	Finally, we collect data on HDDs and CDDs from the NOAA's \citet{CPC} for the same panel. Then, we merge these three data sets and trim 1\% of outliers for a total of $818$ observations for each month and state. We use this preliminary data set to construct the data required for our regressions. That is, each observation in our estimation equation belongs to a set $(t,s,i)$ consisting of two time periods and a state. Hence, we construct each row in our regression data set using unique combinations of $t,s$ where $t \neq s$ for each state $i$. This gives us a total of 6817 observations.  All in all, each observation in our regression data set consists the following variables: state ($i$), date 1 ($t$), date 2 ($s$), the log difference in electricity consumption between month 1 and month 2, the log difference in the price of electricity, the log difference in the price of coal, the number of CDDs for each date, the number of HDDs for each date, and the difference in months between dates 1 and 2.

	\begin{table}[t] \centering 
		\caption{Descriptive Statistics}
		\label{table:stats} 
		\begin{tabular}{@{\extracolsep{0pt}}lrrrrrrr@{}}
			\\[-4ex]\hline  
			\hline \\[-1.8ex] 
			{} &     Mean &      StDev  &     Min &     25\% &      50\% &      75\% &       Max \\
			\midrule
			Electricity Price (\$/kWh)     &     0.12 &     0.03 &    0.08 &    0.10 &     0.11 &     0.13 &      0.21 \\
			Electricity Load (GwH) & 2454.97 &  2380.24 &  144.00 &  790.50 &  1869.50 &  3315.00 &  18621.00 \\
			Coal Price (\$/ton) &    46.00 &    17.68 &   24.04 &   33.43 &    41.40 &    51.83 &    107.49 \\
			Cooling Degree Days      &    64.66 &   134.97 &    0.00 &    0.00 &     0.00 &    47.25 &    761.00 \\
			Heating Degree Days       &   443.20 &   412.27 &    0.00 &   48.00 &   367.00 &   754.50 &   1794.00 \\ [0.5ex]
			\hline 
			\hline \\[-1.8ex] 
			\multicolumn{8}{@{}p{41.5em}@{}}{\textit{Note: } The underlying data covers all 50 US states from 2011 to 2018; outliers are removed by trimming 1\% of each variable. Each observation corresponds to a particular year, month, and state. }  \\ 
		\end{tabular}
	\end{table}

@@

	\begin{table}[t] \centering 
		\caption{OLS Regression Results}
		\label{table:ols} 
		\small
		\begin{tabular}{@{\extracolsep{5pt}}lcccccc} 
			\\[-4ex]\hline  
			\hline \\[-1.8ex] 
			& \multicolumn{6}{c}{\textit{Dependent variable:} $\ln (Z_{ t, i} / Z_{ s, i})$} \\ [0.5ex]
			\cline{2-7} 
			\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\ [0.5ex]
			\hline \\[-1.8ex] 
			$-\ln (P_{t,i} / P_{s,i})$ & 0.937$^{***}$ & 1.075$^{***}$ & 1.098$^{***}$ & 1.074$^{***}$ & 1.263$^{***}$ & 1.305$^{***}$ \\ 
			& (0.030) & (0.026) & (0.026) & (0.224) & (0.172) & (0.169) \\ 
			& & & & & & \\ 
			$\Delta_{t,s}$ &  &  & 0.0005$^{***}$ &  &  & 0.0065$^{*}$ \\ 
			&  &  & (0.0001) &  &  & (0.0003) \\ 
			& & & & & & \\ 
			CDD$_t$ &  & 1.156$^{***}$ & 1.163$^{***}$ &  & 1.164$^{***}$ & 1.174$^{***}$ \\

@@

@@

@@

	\begin{table}[h] 
		\centering 
		\caption{IV (2SLS) Regression Results}
		\label{table:iv} 
		\small
		\begin{tabular}{@{\extracolsep{4pt}}lcccccc} 
			\\[-4ex]\hline  
			\hline \\[-1.6ex] 
			& \multicolumn{3}{c}{First-Stage} & \multicolumn{3}{c}{Second-Stage} \\ [0.5ex]
			& \multicolumn{3}{c}{\textit{Dep. Variable:} $\ln (P_{t,i} / P_{s,i})$ } & \multicolumn{3}{c}{\textit{Dep. Variable:}  $\ln (Z_{ t, i} / Z_{ s, i})$}\\ [0.5ex]
			\cmidrule(lr){2-4} \cmidrule(lr){5-7}\\[-2.2ex] 
			& (A.1) & (B.1) & (C.1) & (A.2) & (B.2) & (C.2)\\ [0.5ex]
			\hline \\[-1.8ex] 
			$ \ln (C_{t,i} / C_{s,i})$ & $-$0.042$^{***}$ & $-$0.018$^{***}$ & $-$0.018$^{***}$ &  &  &  \\ 
			& (0.002) & (0.002) & (0.002) &  &  &  \\ 
			& & & & & & \\ 
			$-\ln (P_{t,i} / P_{s,i})$ &  &  &  & 2.978$^{***}$ & 5.896$^{***}$ & 5.818$^{***}$ \\ 
			&  &  &  & (0.180) & (0.548) & (0.524) \\ 
			& & & & & & \\ 
			$\Delta_{t,s}$  &  &  & 0.001$^{***}$ &  &  & 0.003$^{***}$ \\ 
			&  &  & (0.00004) &  &  & (0.0004) \\ 
			& & & & & & \\ 
			CDD$_t$  &  & 0.100$^{***}$ & 0.105$^{***}$ &  & 1.637$^{***}$ & 1.657$^{***}$ \\ 
			$\quad(\times 1000^{-1})$ &  & (0.006) & (0.006) &  & (0.068) & (0.067) \\ 
			& & & & & & \\ 
			CDD$_s$ &  & $-$0.096$^{***}$ & $-$0.114$^{***}$ &  & $-$1.688$^{***}$ & $-$1.783$^{***}$ \\ 
			$\quad(\times 1000^{-1})$  &  & (0.009) & (0.009) &  & (0.079) & (0.084) \\ 
			& & & & & & \\ 
			HDD$_t$  &  & $-$0.048$^{***}$ & $-$0.048$^{***}$ &  & 0.001 & 0.007 \\ 
			$\quad(\times 1000^{-1})$ &  & (0.003) & (0.003) &  & (0.031) & (0.030) \\ 
			& & & & & & \\ 
			HDD$_s$  &  & 0.053$^{***}$ & 0.055$^{***}$ &  & 0.0001 & 0.007 \\ 
			$\quad(\times 1000^{-1})$  &  & (0.003) & (0.003) &  & (0.035) & (0.034) \\   [0.9ex]
			\hline \\[-1.8ex] 
			State FEs &  Yes & Yes  & Yes  & Yes & Yes & Yes \\ 
			Observations & 6817 & 6817 & 6817 & 6817 & 6817 & 6817 \\ 
			R$^{2}$ & 0.061 & 0.264 & 0.294 &  &  & \\ 
			Adjusted R$^{2}$ & 0.061 & 0.264 & 0.293 & &  &  \\ 
			F Statistic & 443$^{***}$  & 489$^{***}$ & 472$^{***}$  &  &  &  \\ 
			\hline 
			\hline \\[-1.8ex] 
			\multicolumn{7}{@{}p{41.5em}@{}}{\textit{Note: } The log difference in coal price between period $t$ and $s$, $ \ln (C_{t,i} / C_{s,i})$, is used as an instrument in these regressions. The sample covers all 50 US states from 2011 to 2018; outliers are removed by trimming 1\% of each variable except $\Delta_{t,s}$. The unit of observation is a set $(t,s,i)$ where $t \neq s$ are months and $i$ is a state. The coefficient on $\ln (P_{t,i} / P_{s,i})$ is an estimate of $-\sigma$. The variable $\Delta_{t,s}$ is the difference in months between periods $t$ and $s$. CDD$_t$ and HDD$_t$ refer to the total number of heating and cooling degree days in month $t$. We scale the coefficients on degree days for clarity. Robust standard errors are reported in parentheses. *p$\textless$0.05, **p$\textless$0.01, ***p$\textless$0.001}  \\ 
		\end{tabular} 
	\end{table}

	\begin{table}[!t] 
		\centering 
		\caption{Partially Linear IV Regression Results}
		\label{table:pariv} 
		\small
		\begin{tabular}{@{\extracolsep{5.3em}}lccc} 
			\\[-4ex]\hline  
			\hline \\[-1.8ex] 
			& \multicolumn{3}{c}{\textit{Instrument: $ \ln (C_{t,i} / C_{s,i})$}} \\ 
			\cline{2-4} 
			%\\[-1.8ex] & ln\_load\_rel \textasciitilde ln\_price\_rel & ln\_load\_rel \textasciitilde (CDD\_1) & ln\_load\_rel \textasciitilde time\_diff + (CDD\_1) \\ 
			\\[-1.8ex] & (1) & (2) & (3)\\ [0.5ex]
			\hline \\[-1.8ex] 
			$\hat{\sigma} $ & 2.9976$^{***}$ & 1.2123$^{***}$ & 0.8847$^{***}$ \\ 
			& (0.169) & (0.052) & (0.044) \\ [0.9ex]
			\hline \\[-1.8ex] 
			Time Control &   &   & Yes  \\ 
			Degree Day Controls &   & Yes  & Yes  \\ 
			Observations & 6817 & 6817 & 6817 \\
			\hline 
			\hline \\[-1.8ex] 
			\multicolumn{4}{@{}p{41.5em}@{}}{\textit{Note: } The log difference in coal price between period $t$ and $s$, $ \ln (C_{t,i} / C_{s,i})$, is used as an instrument in these regressions. The sample covers all 50 US states from 2011 to 2018; outliers are removed by trimming 1\% of each variable except $\Delta_{t,s}$. The unit of observation is a set $(t,s,i)$ where $t \neq s$ are months and $i$ is a state. The estimation procedure is described in the appendix.  Robust standard errors are reported in parentheses. *p$\textless$0.05, **p$\textless$0.01, ***p$\textless$0.001}  \\ 
		\end{tabular} 
	\end{table}

@@

	\begin{center}
		\noindent \textbf{Example  A:} $\alpha_t = 0.6$, $\alpha_s = 0.4$, $\xi_1 = (1, \, 1)$, $\xi_2 = (1, \, 0.1)$, $c_1 = 104.3$, $c_2 = 60$. 
		\label{Params: Example A}
	\end{center}

@@

	\begin{figure}[t]
		\caption{The Elasticity of Substitution between Solar and Coal} 
		\label{fig:eosnum}
		\footnotesize
		\vspace{-1em}
		\begin{tabular}{@{\extracolsep{0em}}c}
			\includegraphics[width=1\linewidth]{../figures/fig_elasticity} \\
			\multicolumn{1}{@{\hspace{0.2in}}p{5.9in}@{}}{ \textit{Note: } Technology 1 is coal and technology 2 is solar. The legend in the upper subplot also applies to the lower subplot. These results were obtained using the following parameters: $\alpha_t = 0.6$, $\alpha_s = 0.4$, $\xi_1 = (1, \, 1)$, $\xi_2 = (1, \, 0.1)$, $c_1 = 104.3$, $c_2 = 60$. Furthermore, we set the parameter for the intertemporal elasticity of substitution for electricity consumption equal to our estimate $\hat{\sigma} = 0.8847$. In order to generate these numerical results, we first found the optimal quantities of $X$ over a range of prices $c_1^* \in (0.5\, c_1, 2 \,c_1)$. Then, we obtained estimates of the elasticity of substitution by numerically differentiating $\ln(X_1/X_2)$ with respect to $-\ln(c_1/ c_2)$. That is, the elasticity of substitution  between technology 1 and 2 is given by the slope of the upper subplot, and it is graphed in the lower subplot. Finally, we repeat this procedure with $\sigma$ equal to two standard deviations above and below its estimated value $\hat{\sigma}$; that is, the dashed lines represent  $\sigma = 0.8847 \pm (1.96)(0.044)$. }     \\  
		\end{tabular}
	\end{figure}

@@

	\begin{figure}[!t]
		\caption{The Price Elasticity of Demand for Coal Power} 
		\label{fig:coalelas}
		\footnotesize
		\vspace{-1em}
		\begin{tabular}{@{\extracolsep{0em}}c}
			\includegraphics[width=1\linewidth]{../figures/fig_coal_elas} \\
			\multicolumn{1}{@{\hspace{0.2in}}p{5.9in}@{}}{ \textit{Note: } These results were obtained using the following parameters: $\alpha_t = 0.6$, $\alpha_s = 0.4$, $\xi_1 = (1, \, 1)$, $\xi_2 = (1, \, 0.1)$, $c_1 = 104.3$, $c_2 = 60$, $\sigma = 0.5$. We generate these results by finding the optimal quantity of coal, $X_1$, over a range of percent changes in its price $c_1$. Then, on the y-axis, we plot the log difference in $X_1$ divided by the log difference in its price. This is equivalent to the price elasticity of demand for $X_1$. }     \\  
		\end{tabular}
	\end{figure}

@@

	\begin{figure}[t]
		\captionsetup{format=centerproper}
		\caption{The Effect of Battery Storage on the Elasticity\newline of Substitution between Solar and Coal} 
		\label{fig:battery}
		\footnotesize
		\vspace{-1em}
		\begin{tabular}{@{\extracolsep{0em}}c}
			\includegraphics[width=1\linewidth]{../figures/fig_batteries} \\
			\multicolumn{1}{@{\hspace{0.2in}}p{5.9in}@{}}{ \textit{Note: } Technology 1 is coal and technology 2 is solar. The legend in the upper subplot also applies to the lower subplot. The elasticity of substitution  between technology 1 and 2 is given by the slope of the upper subplot, and it is graphed in the lower subplot.  These results were obtained using the following parameters: $\alpha_t = 0.6$, $\alpha_s = 0.4$, $\xi_1 = (1, \, 1)$, $\xi_2 = (1, \, 0.1)$, $c_1 = 104.3$, $c_2 = 60$. Furthermore, we set the parameter for the intertemporal elasticity of substitution for electricity consumption equal to our estimate $\hat{\sigma} = 0.8847$. We generated these numerical results with the same procedure used for \autoref{fig:eosnum}. We repeated this procedure with $\xi_2 = (0.95, 0.15)$ and $\xi_2 = (0.90, 0.20)$ to simulate the effects of shifting solar power output using batteries. }     \\  
		\end{tabular}
	\end{figure}

@@

	\begin{table}[!t] \centering 
		\captionsetup{format=centerproper}
		\caption{The Effect of Battery Storage on the   \newline Elasticity of Demand for Coal Power} 
		\label{table:storelas} 
		\small
		\begin{tabular}{@{\extracolsep{2em}}ccc} 
			\\[-4ex]\hline  
			\hline \\[-1.8ex] 
			\\[-2.6ex] \specialcell{Solar Output during \\ Peak Hours ($\xi_{2t}$)} &  \specialcell{Solar Output during \\ Off-Peak Hours ($\xi_{2s}$)}  & \specialcell{The Elasticity of  \\ Demand for Coal Power}\\ [0.5ex]
			\hline \\[-1.8ex] 
			100\% &  \phantom{1}5\%  &  -3.25 \\
			\phantom{1}95\% &            10\%  &  -3.71 \\
			\phantom{1}90\% &            15\%  &  -4.40 \\
			\phantom{1}85\% &            20\%  &  -5.48 \\
			\phantom{1}80\% &            25\%  &  -7.29 \\
			\hline 
			\hline \\[-1.8ex] 
			\multicolumn{3}{@{}p{34.2em}@{}}{\textit{Note: } These results were obtained using the following parameters: $\alpha_t = 0.6$, $\alpha_s = 0.4$, $\xi_1 = (1, \, 1)$, $c_1 = 104.3$, $c_2 = 60$, $\sigma = 0.8847$. We generated these results by numerically differentiating the optimal quantity of coal power, $X_1$, with respect to its own price, $c_1$, around its initial price $104.3$. We repeated this process for various values of the output parameter for solar power, $\xi_{2}$. Specifically, we considered 5\% to 20\% shifts in the output of solar power from peak to  off-peak hours to simulate the effects of implementing battery storage. }  \\ 
		\end{tabular} 
	\end{table}

@@

>>>>

\documentclass[11pt,a4paper,leqno]{extarticle}
\usepackage[margin=1in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{booktabs} % for toprule, midrule and bottomrule
\usepackage{adjustbox}
\usepackage{amsmath}
\usepackage{bbold}
\usepackage{etoolbox}
\usepackage{setspace} % for \onehalfspacing and \singlespacing macros
\usepackage[hidelinks]{hyperref}
\usepackage{array}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{caption}
\usepackage{pdflscape}
\usepackage{caption}
\usepackage{tabularx}
\usepackage{authblk}
\usepackage{float}
\usepackage{siunitx}
\usepackage{titlesec}
\usepackage{pgfplots}
\usepackage[authoryear]{natbib}
\usepackage{scrextend}
\usepackage{nicefrac}
\usepackage{enumitem}
\usepackage{multirow}
\usepackage{xcolor}
\usepackage{cleveref}
\usepackage{varwidth}
\usepackage{gensymb}

@@

\titleformat{\section}
{\bf\centering\large}{\thesection}{1em}{}
\titleformat{\subsection}
{\itshape\centering}{\thesubsection}{1em}{}
\titleformat{\subsubsection}
{\bf}{\thesubsubsection}{1em}{}

% images
\graphicspath{ {D:/Users/saketh/Documents/GitHub/Energy-Intermittency-Paper/figures/} }

@@

@@

@@

	\title{\singlespacing{\textbf{Efficient Pollution Abatement in Electricity Markets with Intermittent Renewable Energy}}\thanks{Saketh Aleti: Graduate Student, Department of Economics, Duke University, 419 Chapel Drive, 213 Social Sciences Bldg., Durham, NC 27708-0097, USA (email: sa400@duke.edu). Gal Hochman: Associate Professor, Department of Agriculture, Food \& Resource Economics, Cook Office Building, 55 Dudley Road, Room 116, New Brunswick, NJ 08901, USA (email: gal.hochman@rutgers.edu).}}

	\author[]{Saketh Aleti and Gal Hochman}

@@

	% Syntax: \begin{addmargin}[<left indentation>]{<indentation>}
	\begin{addmargin}[0.5in]{0.5in}
		\textit{In this paper, we present a model of the electricity sector where generation technologies are intermittent. The economic value of a electricity generation technology is given by integrating its production profile with the market price of electricity. We parametrize this model empirically and generate numerical results. The introduction of intermittency results in a non-constant elasticity of substitution between renewable and fossil energy. Consequently, this suggests that the efficacy and welfare effects of carbon taxes and renewable subsidies vary geographically. Subsidizing research into battery technology can mitigate this distributional side effect while complementing traditional policies used to promote renewable energy. 
		}
		\\
		\noindent\textbf{Key words:} renewable energy, intermittency, pollution, environment

		\noindent\textbf{JEL Classifications:} Q28, Q41, Q48, Q52, Q55, Q58
	\end{addmargin}

@@

	Consumers purchase a quantity of electricity $Z_t$ in each period $t$. Furthermore, they demand a greater quantity of electricity in certain periods; for instance, consumers need more electricity during the middle of the day more than at night. At the same time, consumers are willing to shift their consumption from one period to another in response to a shift in prices. Overall, these characteristics can be captured using a standard CES utility function.  Furthermore, since CES preferences are homothetic, we may aggregate the consumers into a single representative consumer. This consumer has the utility function
	\begin{equation}
	U = \left( \sum_t \alpha_t Z_t^\phi  \right)^{1/\phi}
	\end{equation}
	where $\sigma = 1/(1-\phi)$ is the intertemporal elasticity of substitution for electricity consumption. To ensure that indifferences surfaces are convex, we have $\phi \in (0,1)$ or, equivalently,  $\sigma > 0$; additionally, since electricity consumption increases utility, we must have $\alpha_t > 0$ for all $t$.  For simplicity, we also define $\sum_t \alpha_t = 1$, so that a 1\% increase in electricity consumption in all periods increases utility by 1\%; consequently, we then have $\alpha_t \in (0,1)$. The budget constraint is given by
	\begin{equation}
	I = \sum_t p_t Z_t
	\end{equation}
	where $p_t$ is the price of electricity in period $t$ and $I$ is the income. Our representative consumer maximizes utility against this budget constraint; the first order conditions of this problem imply:
	\begin{align}\label{eq:demand}
	Z_t &= \left(\frac{\alpha_t}{p_t} \right)^\sigma \frac{I}{P} \\
	P &= \sum_t \alpha_t^\sigma p_t^{1-\sigma}
	\end{align}
	where $P$ is the price index. Note that this model naturally does not allow for blackouts in equilibrium, since the price of electricity in any period gets arbitrarily large as the quantity of energy consumed in that period approaches 0. Furthermore, note that prices must be positive; although this is sometimes violated in reality, we do not believe that this assumption significantly affects our analysis.

@@

	Next, our representative firm sets $X$ once to maximize total profit while facing the cost function $C(X_1, \dots, X_n)$.  \citet{HH} note that past literature has argued for concave cost functions in the energy market due to effects such as economies of scale and learning-by-doing; on the other hand, standard cost functions are generally convex. So, like Helm and Meir, we take an intermediate approach by using a linear cost function. Specifically, we have $C(X) \equiv \sum_i c_i X_i \equiv c^T X$ where $c_i$ is the cost per unit of $X_i$. Total profit is given by 
	\begin{equation}
	\Pi = p^T Z - c^T X
	\end{equation}
	To simplify the algebra, we set the number of technologies equal to the number of periods $(n=m)$. Additionally, we further require that the output per unit of each technology is unique and non-negative in each period; in other words, the output per unit of one technology is not a linear combination of those of the other technologies in our set. This then implies that $\xi$ is of full rank and therefore invertible.  Now, maximizing profit, we find the first order condition:
	\begin{equation}
	\frac{\partial \Pi}{\partial X} = 0 \implies p = \xi^{-1} c
	\end{equation}
	Combining this FOC with the demand equation (\cref{eq:demand}) allows us to find the equilibrium. Generally, the equilibrium results for any number of technologies ($n$) are analytic, but they are difficult to interpret due to the number of parameters involved.

@@

	Now, recall the demand equation from earlier
	$$Z_t = \left(\frac{\alpha_t}{p_t} \right)^\sigma \frac{I}{P}$$
	where $P$ is the price index and $I$ is income. Retail customers pay fixed rates each month for electricity, hence $p_t$ is constant within each month; and, we expect that income $I$ does not vary significantly on a daily basis. Consequently, all variation in intramonthly demand is due to the share parameter $\alpha$ and the elasticity $\sigma$. But, this creates a problem; since retail consumers do not face prices that vary each hour, we cannot estimate  $\sigma$ on an hourly basis.

@@

	We begin with the demand equation from our general model:
	\begin{align*}
	Z_t &= \left(\frac{\alpha_t}{p_t} \right)^\sigma \frac{I}{P} \\
	P &= \sum_t \alpha_t^\sigma p_t^{1-\sigma}
	\end{align*}
	For any pair of electricity outputs $Z_t$ and $Z_s$, we have:
	$$\frac{Z_t}{Z_s} = \left(\frac{\alpha_t \, p_s}{\alpha_s \, p_t} \right)^\sigma \\$$
	Taking logs on both sides and letting $i$ represent different observations, we may rewrite this in a form more suitable for estimation. 
	\begin{align*}
	\ln (Z_{t, i} / Z_{ s, i}) &= -\sigma \ln (P_{t,i} / P_{s,i}) + \sigma \ln (\alpha_{t,i} / \alpha_{s,i}) 
	\end{align*}
	Our data differentiates consumption for each state in the US, so we let $i$ refer to a particular state. Additionally, most consumers pay monthly fixed rates for electricity, so we can, at most, estimate this equation on a monthly basis; hence, $t$ and $s$ refer to different months. Also, note that each observation only corresponds to a single state $i$; this is because consumers within each state can substitute consumption across time, but consumers in different states do not substitute consumption with one another.

@@

	We collect monthly data from the \cite{EIANetgen} on retail electricity prices and consumption for each state in the US from 2011 to 2018. Additionally, also from the EIA, we obtain data on the average cost of coal for electricity generation for each state and month.\footnote{ The coal price data set contains a large number of missing values due to privacy concerns; however, we do not expect that these missing values are correlated with the data itself.} We deflated both electricity and coal prices over time using the PCEP Index provided by the  \cite{USBEA}. Finally, we collect data on HDDs and CDDs from the NOAA's \citet{CPC} for the same panel. Then, we merge these three data sets and trim 1\% of outliers for a total of $818$ observations for each month and state. We provide descriptive statistics for this data in \autoref{table:stats}. We use this preliminary data set to construct the data required for our regressions. That is, each observation in our estimation equation belongs to a set $(t,s,i)$ consisting of two time periods and a state. Hence, we construct each row in our regression data set using unique combinations of $t,s$ where $t \neq s$ for each state $i$. This gives us a total of 6817 observations.  All in all, each observation in our regression data set consists the following variables: state ($i$), date 1 ($t$), date 2 ($s$), the log difference in electricity consumption between month 1 and month 2, the log difference in the price of electricity, the log difference in the price of coal, the number of CDDs for each date, the number of HDDs for each date, and the difference in months between dates 1 and 2.

	\vspace{0.15in}
	\begin{center}
		[INSERT Table 1: Descriptive Statistics]
	\end{center}
	\vspace{0.15in}

@@

@@

	\vspace{0.15in}
	\begin{center}
		[INSERT Table 2: OLS Regression Results]
	\end{center}
	\vspace{0.15in}

@@

	\vspace{0.15in}
	\begin{center}
		[INSERT Table 3: IV (2SLS) Regression Results]
	\end{center}
	\vspace{0.15in}

@@

	\vspace{0.15in}
	\begin{center}
		[INSERT Table 4: Partially Linear IV Regression Results]
	\end{center}
	\vspace{0.15in}

@@

	\begin{center}
		\noindent \textbf{Model Parameters:} $\alpha_t = 0.6$, $\alpha_s = 0.4$, $\xi_1 = (1, \, 1)$, $\xi_2 = (1, \, 0.1)$, $c_1 = 104.3$, $c_2 = 60$. 
		\label{Params: Example A}
	\end{center}

@@

	\vspace{0.15in}
	\begin{center}
		[INSERT Figure 1: The Elasticity of Substitution between Solar and Coal]
	\end{center}
	\vspace{0.15in}

@@

	\vspace{0.15in}
	\begin{center}
		[INSERT Figure 2: The Price Elasticity of Demand for Coal Power]
	\end{center}
	\vspace{0.15in}

@@

	\vspace{0.15in}
	\begin{center}
		[INSERT Figure 3: The Effect of Battery Storage on the Elasticity of Substitution between Solar and Coal]
	\end{center}
	\vspace{0.15in}

@@

	\vspace{0.15in}
	\begin{center}
		[INSERT Figure 4: The Effect of Battery Storage on the  Elasticity of Demand for Coal Power]
	\end{center}
	\vspace{0.15in}

@@

	\section{Acknowledgments}

	This research was funded by the Sun Grant (122682). We would also like to thank Prof. Roger Klein and the participants of the NAREA 2019 Workshop for helpful comments.

	\clearpage

	\section{Tables and Figures}
